---
title: "Portofolio Case Study, Divvy"
author: "Freid Kholy"
date: "2023-02"
output: html_document
---

### Install the packages you need for this job.

```{r}
install.packages("tidyverse")
install.packages("lubridate")
install.packages("janitor")
install.packages("skimr")
```

### Call the librarys

```{r}
library(tidyverse)
library(lubridate)
library(janitor)
library(skimr)
```

### Import the data and rename them whith simple names.

```{r Import and rename}
trips2004 <- read.csv('202004-divvy-tripdata.csv')
trips2005 <- read.csv('202005-divvy-tripdata.csv')
trips2006 <- read.csv('202006-divvy-tripdata.csv')
trips2007 <- read.csv('202007-divvy-tripdata.csv')
trips2008 <- read.csv('202008-divvy-tripdata.csv')
trips2009 <- read.csv('202008-divvy-tripdata.csv')
trips2010 <- read.csv('202010-divvy-tripdata.csv')
trips2011 <- read.csv('202011-divvy-tripdata.csv')
trips2012 <- read.csv('202012-divvy-tripdata.csv')
trips2101 <- read.csv('202101-divvy-tripdata.csv')
trips2102 <- read.csv('202102-divvy-tripdata.csv')
trips2103 <- read.csv('202103-divvy-tripdata.csv')
trips2104 <- read.csv('202104-divvy-tripdata.csv')
trips2105 <- read.csv('202105-divvy-tripdata.csv')
trips2106 <- read.csv('202106-divvy-tripdata.csv')
trips2107 <- read.csv('202107-divvy-tripdata.csv')
trips2108 <- read.csv('202108-divvy-tripdata.csv')
trips2109 <- read.csv('202109-divvy-tripdata.csv')
trips2110 <- read.csv('202110-divvy-tripdata.csv')
trips2111 <- read.csv('202111-divvy-tripdata.csv')
trips2112 <- read.csv('202112-divvy-tripdata.csv')
trips2201 <- read.csv('202201-divvy-tripdata.csv')
trips2202 <- read.csv('202202-divvy-tripdata.csv')
trips2203 <- read.csv('202203-divvy-tripdata.csv')
trips2204 <- read.csv('202204-divvy-tripdata.csv')
```

#### Check the column names.

Taking a look at the elements of the individual table,
before merging individual files into one integrated file

Note: the example shows only one, April 2020 file, but a full examination of each files column headers should be pursued

```{r check the column names}
colnames(trips2004)
```

#### Take a look at the data table.

Taking a look at the elements of the individual table,
before merging individual files into one integrated file

Note: the example shows only one, April 2020 file, but a full examination of each files table details should be pursued

```{r look over the table}
View(trips2004)
```

#### Check the data format alongside to the column names of each column

Taking a look at the elements of the individual table,
before merging individual files into one integrated file

Note: the example shows only one, April 2020 file, but a full examination of each files column data format should be pursued

```{r column names and data format}
str(trips2004)
```

### Compare_df_cols.

This command let us know if columns will successfully bind together by rows.

As you can see from the results below, We can see that column end_station_id and column start_station_id have different data formats in several columns. From April to November 2020, it is integer, and after December, it is character.
Integration is possible only when those are changed to the same data format.

```{r compare_df_cols}
compare_df_cols(trips2004, trips2005, trips2006, trips2007, trips2008, trips2009, trips2010, trips2011, trips2012, trips2101, trips2102, trips2103, trips2104, trips2105, trips2106, trips2107, trips2108, trips2109, trips2110, trips2111, trips2112, trips2201, trips2202, trips2203, trips2204, return = "mismatch")
```

### Change the integer data to character format.

If the data format is different, integration into one file is impossible, so make the data format matched.

The targeted columns are the end_station_id and end_station_id from April to November.

```{r changing the data fotrmat}
trips2004 <- mutate(trips2004, end_station_id = as.character(end_station_id), start_station_id = as.character(start_station_id))
trips2005 <- mutate(trips2005, end_station_id = as.character(end_station_id), start_station_id = as.character(start_station_id))
trips2006 <- mutate(trips2006, end_station_id = as.character(end_station_id), start_station_id = as.character(start_station_id))
trips2007 <- mutate(trips2007, end_station_id = as.character(end_station_id), start_station_id = as.character(start_station_id))
trips2008 <- mutate(trips2008, end_station_id = as.character(end_station_id), start_station_id = as.character(start_station_id))
trips2009 <- mutate(trips2009, end_station_id = as.character(end_station_id), start_station_id = as.character(start_station_id))
trips2010 <- mutate(trips2010, end_station_id = as.character(end_station_id), start_station_id = as.character(start_station_id))
trips2011 <- mutate(trips2011, end_station_id = as.character(end_station_id), start_station_id = as.character(start_station_id))
```

#### Check if the data format has changed correctly.

A result of 0_rows means that there is no mismatched data format anymore.

```{r confirming 0 mismatchs}
compare_df_cols(trips2004, trips2005, trips2006, trips2007, trips2008, trips2009, trips2010, trips2011, trips2012, trips2101, trips2102, trips2103, trips2104, trips2105, trips2106, trips2107, trips2108, trips2109, trips2110, trips2111, trips2112, trips2201, trips2202, trips2203, trips2204, return = "mismatch")
```

#### Checking if the data format has changed correctly.

As shown in the results below, it can be seen that the data format of the two columns we targeted has been well changed from integer to character.


```{r}
str(trips2004)
```

### Cobining tables into one integrated file.

Finally!! lets integrate several tables into one integrated file.

```{r}
trips2004_2204 <- bind_rows(trips2004, trips2005, trips2006, trips2007, trips2008, trips2009, trips2010, trips2011, trips2012, trips2101, trips2102, trips2103, trips2104, trips2105, trips2106, trips2107, trips2108, trips2109, trips2110, trips2111, trips2112, trips2201, trips2202, trips2203, trips2204)
```

#### Check the integrated file.

Lets see if it is well integrated.

look over the integrated table. The results show a huge single table file with all data integrated from April 2020 to April 2022.

```{r}
View(trips2004_2204)
```

### Character to Date-time conversion.

The integration went well and now we have to analyze it.
However, the data format of the travel start time started_at and the travel end time ended_at; still in the format of character, not in the format of date-time format and should be changed.
We use 'strptime' function for Date-time Conversion from Character

```{r From character to Date-time conversion}
trips2004_2204$started_at = strptime(trips2004_2204$started_at, format = "%Y-%m-%d %H:%M:%S")
trips2004_2204$ended_at = strptime(trips2004_2204$ended_at, format = "%Y-%m-%d %H:%M:%S")
```

#### See if the data format has changed well.

As you can see, the data format of the target two columns was well changed from character to date-time format.
If you can check the POSIXlt format, it is done properly.

```{r}
str(trips2004_2204)
```

#### The total number of rows and columns.

```{r}
dim(trips2004_2204)
```

#### Touch the outline with the summary function.

```{r}
summary(trips2004_2204)
```

### Create new columns: "month, year, and day of the week for analysis".

Now, for analysis by month, year, and day of the week, extract separately from the started_at column. We first extract dates only from the started_at column to create new date columns, and then extract months, day, and day of the week.

```{r Creating Date, Month, Day, Day of the week colunms}
trips2004_2204$date <- as.Date(trips2004_2204$started_at)
trips2004_2204$year <- format(as.Date(trips2004_2204$date), "%Y")
trips2004_2204$month <- format(as.Date(trips2004_2204$date), "%m")
trips2004_2204$day <- format(as.Date(trips2004_2204$date), "%d")
trips2004_2204$day_of_week <- format(as.Date(trips2004_2204$date), "%A")
```

### Find the length of each tirp.

Create a ride_length column through summing the difference between ended_at & started_at columns using the 'difftime' function. 

```{r ride length}
trips2004_2204$ride_length <- difftime(trips2004_2204$ended_at, trips2004_2204$started_at)
```

#### The current work status.

Lets check if the work so far has gone well.

```{r}
View(trips2004_2204)
str(trips2004_2204)
```

### Let's change the data format of ride_lengh to 'numeric'.

We need to change the format to numeric to analyze and calculate.

When checked using the str function, it changed to numeric.

```{r}
trips2004_2204$ride_length <- as.numeric(as.character(trips2004_2204$ride_length))

str(trips2004_2204)
```

### Remove unnecessary values.

remove some unnecessary values: rides with ride_length less than a Zero.

```{r}
trips2004_2204 <- trips2004_2204[!(trips2004_2204$ride_length<0),]
```

#### Final inspection.

A final checking before extracting the organized and cleaned integrated file.

NA values can be deleted from R or simply excluded from Tableau. I did not work on R because exclusion from Tableau is simpler.

```{r}
summary(trips2004_2204)
summary(trips2004_2204$ride_length)
```
### Make a sample from the dataset

Extract a sample of the data set for performing an easy analysis 
An estimate for the best size of the sample was a 37% of the total integrated file which is equaled to 3,490,668 rows.

```{r}
F <- sample(1:nrow(trips2004_2204), 3490668)
trips2004_2204_sample <- trips2004_2204[F,]
```


### File Extraction.

Finally! extract the original integrated file for backup purposes and the sample file for performing the analysis in Tableau.

```{r}
write.csv(trips2004_2204, "trips2004_2204.csv")
write(trips2004_2204_sample,"trips2004_2204_sample.csv")
```