---
title: "Waleed Said Forecasted Data R Markdown"
author: "Fareid El Kholy"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(repos = c(CRAN = "https://cran.r-project.org"))
```

## R Markdown

### installing packages

```{r}
install.packages("readxl")
install.packages("TSA")
install.packages("TTR")
install.packages("tseries")
install.packages("urca")
```

### Installing the "AnalyzTS" Package manully

```{r}
# The installation needs to be manually from the site due to it being removed from the online server, and for the path of the local file to be specified manually
install.packages("C:/Users/Dr.Haba Al-Ghonaimy/Downloads/AnalyzeTS_2.3.tar.gz",
                 repos = NULL, type = "source")
```

### loading needed packages

```{r}
library(readxl)
library(urca)
library(TSA)
library(TTR)
library(tseries)
library(AnalyzeTS)
```

### Importing the data-set

```{r}
# The path would be a different one depending on the user computer 
file_path <- "C:/Users/Dr.Haba Al-Ghonaimy/Downloads/before.xlsx"
Waleed_Data <- read_excel(file_path)
```

### Converting the 'Years' column to a Date format column

```{r}
Waleed_Data$Years <- as.Date(paste(Waleed_Data$Years, "-01-01", sep = ""), 
                                     format = "%Y-%m-%d")
str(Waleed_Data)
```

### Making every XB Time Series

```{r}
XB1_ts <- ts(Waleed_Data$XB1, frequency = 1)
XB2_ts <- ts(Waleed_Data$XB2, frequency = 1)
XB3_ts <- ts(Waleed_Data$XB3, frequency = 1)
XB4_ts <- ts(Waleed_Data$XB4, frequency = 1)
XB5_ts <- ts(Waleed_Data$XB5, frequency = 1)
```

### "Mamedova" & "FVD" models applications:

##### XB1

```{r}
XB1_MForcast <- fuzzy.ts2(XB1_ts, n = 7, w = 7, D1 = 0, D2 = 0, C = 0.0001, forecast = 12, 
                  r = 12, trace = FALSE, plot = TRUE, grid = FALSE, type = "Abbasov-Mamedova")
XB1_NForcast <- fuzzy.ts2(XB1_ts, n = 7, w = 7, D1 = 0, D2 = 0, C = 0.0001, forecast = 12, 
                          r = 12, trace = FALSE, plot = TRUE, grid = FALSE, type = "NFTS")
```

##### XB2

```{r}
XB2_MForcast <- fuzzy.ts2(XB2_ts, n = 7, w = 7, D1 = 0, D2 = 0, C = 0.0001, forecast = 12, 
                  r = 12, trace = FALSE, plot = TRUE, grid = FALSE, type = "Abbasov-Mamedova")
XB2_NForcast <- fuzzy.ts2(XB2_ts, n = 7, w = 7, D1 = 0, D2 = 0, C = 0.0001, forecast = 12, 
                          r = 12, trace = FALSE, plot = TRUE, grid = FALSE, type = "NFTS")

```

##### XB3

```{r}
XB3_MForcast <- fuzzy.ts2(XB3_ts, n = 7, w = 7, D1 = 0, D2 = 0, C = 0.0001, forecast = 12, 
                  r = 12, trace = FALSE, plot = TRUE, grid = FALSE, type = "Abbasov-Mamedova")
XB3_NForcast <- fuzzy.ts2(XB3_ts, n = 7, w = 7, D1 = 0, D2 = 0, C = 0.0001, forecast = 12, 
                          r = 12, trace = FALSE, plot = TRUE, grid = FALSE, type = "NFTS")

```

##### XB4

```{r}
XB4_MForcast <- fuzzy.ts2(XB4_ts, n = 7, w = 7, D1 = 0, D2 = 0, C = 0.0001, forecast = 12, 
                  r = 12, trace = FALSE, plot = TRUE, grid = FALSE, type = "Abbasov-Mamedova")
XB4_NForcast <- fuzzy.ts2(XB4_ts, n = 7, w = 7, D1 = 0, D2 = 0, C = 0.0001, forecast = 12, 
                          r = 12, trace = FALSE, plot = TRUE, grid = FALSE, type = "NFTS")

```

##### XB5

```{r}
XB5_MForcast <- fuzzy.ts2(XB5_ts, n = 7, w = 7, D1 = 0, D2 = 0, C = 0.0001, forecast = 12, 
                  r = 12, trace = FALSE, plot = TRUE, grid = FALSE, type = "Abbasov-Mamedova")
XB5_NForcast <- fuzzy.ts2(XB5_ts, n = 7, w = 7, D1 = 0, D2 = 0, C = 0.0001, forecast = 12, 
                          r = 12, trace = FALSE, plot = TRUE, grid = FALSE, type = "NFTS")

```

### Converting every Time-Series forecasted data into a "Data-Frame" format to preform further analysis

```{r}
# XB1 "Mamedova" Model
XB1_MForcast_df <- as.data.frame(XB1_MForcast)
XB1_MForcast_df <- cbind(XB1_MForcast$interpolate, XB1_MForcast$forecast)
XB1_MForcast_df <- c(XB1_MForcast$interpolate, XB1_MForcast$forecast)
# XB1 "FVD" Model
XB1_NForcast_df <- as.data.frame(XB1_NForcast)
XB1_NForcast_df <- cbind(XB1_NForcast$interpolate, XB1_NForcast$forecast)
XB1_NForcast_df <- c(XB1_NForcast$interpolate, XB1_NForcast$forecast)

# XB2 "Mamedova" Model
XB2_MForcast_df <- as.data.frame(XB2_MForcast)
XB2_MForcast_df <- cbind(XB2_MForcast$interpolate, XB2_MForcast$forecast)
XB2_MForcast_df <- c(XB2_MForcast$interpolate, XB2_MForcast$forecast)
# XB2 "FVD" Model
XB2_NForcast_df <- as.data.frame(XB2_NForcast)
XB2_NForcast_df <- cbind(XB2_NForcast$interpolate, XB2_NForcast$forecast)
XB2_NForcast_df <- c(XB2_NForcast$interpolate, XB2_NForcast$forecast)

# XB3 "Mamedova" Model
XB3_MForcast_df <- as.data.frame(XB3_MForcast)
XB3_MForcast_df <- cbind(XB3_MForcast$interpolate, XB3_MForcast$forecast)
XB3_MForcast_df <- c(XB3_MForcast$interpolate, XB3_MForcast$forecast)
# XB3 "FVD" Model
XB3_NForcast_df <- as.data.frame(XB3_NForcast)
XB3_NForcast_df <- cbind(XB3_NForcast$interpolate, XB3_NForcast$forecast)
XB3_NForcast_df <- c(XB3_NForcast$interpolate, XB3_NForcast$forecast)

# XB4 "Mamedova" Model
XB4_MForcast_df <- as.data.frame(XB4_MForcast)
XB4_MForcast_df <- cbind(XB4_MForcast$interpolate, XB4_MForcast$forecast)
XB4_MForcast_df <- c(XB4_MForcast$interpolate, XB4_MForcast$forecast)
# XB4 "FVD" Model
XB4_NForcast_df <- as.data.frame(XB4_NForcast)
XB4_NForcast_df <- cbind(XB4_NForcast$interpolate, XB4_NForcast$forecast)
XB4_NForcast_df <- c(XB4_NForcast$interpolate, XB4_NForcast$forecast)

# XB5 "Mamedova" Model
XB5_MForcast_df <- as.data.frame(XB5_MForcast)
XB5_MForcast_df <- cbind(XB5_MForcast$interpolate, XB5_MForcast$forecast)
XB5_MForcast_df <- c(XB5_MForcast$interpolate, XB5_MForcast$forecast)
# XB5 "FVD" Model
XB5_NForcast_df <- as.data.frame(XB5_NForcast)
XB5_NForcast_df <- cbind(XB5_NForcast$interpolate, XB5_NForcast$forecast)
XB5_NForcast_df <- c(XB5_NForcast$interpolate, XB5_NForcast$forecast)
```

### Creating the "Date" column

```{r}
# Creating a sequence of dates from 2000-01-01 to 2024-12-31
dates <- seq(as.Date("2000-01-01"), as.Date("2023-12-31"), by = "year")


# Convert the dates into a data frame
dates_df <- data.frame(Date = dates)
```

### Combining all of the forecasted data alongside with the Date column

```{r}
combined_data <- cbind(dates_df,
                       XB1_MForcast_df, XB1_NForcast_df,
                       XB2_MForcast_df, XB2_NForcast_df,
                       XB3_MForcast_df, XB3_NForcast_df,
                       XB4_MForcast_df, XB4_NForcast_df,
                       XB5_MForcast_df, XB5_NForcast_df)

# Renaming the columns 
colnames(combined_data) <- c("Date",
                             "XB1 Mamedova Forecast", "XB1 FVD Forecast",
                             "XB2 Mamedova Forecast", "XB2 FVD Forecast",
                             "XB3 Mamedova Forecast", "XB3 FVD Forecast",
                             "XB4 Mamedova Forecast", "XB4 FVD Forecast",
                             "XB5 Mamedova Forecast", "XB5 FVD Forecast")
```

### Creating the Combined Forecasted Data plot

```{r}
# Calling the needed packages 
library(tidyr)
library(ggplot2)

# Pivoting the data
long_df <- tidyr::pivot_longer(combined_data, cols = -Date, names_to = "Variable", 
                                 values_to = "Value")

# Plotting the data
 ggplot(long_df, aes(x = Date, y = Value, color = Variable)) +
     geom_line() +
     labs(title = "Forecasted Data",
          x = "Years",
          y = "Value") +
     theme_minimal()
 
```

### Combining the forecasted data and old data in one file

```{r}
OG_and_Forecasted_df <- data.frame(Date = combined_data$Date,
                                   XB1 = rep(c(Waleed_Data$XB1, rep(NA, 12))),
XB1_Mamedova_Forecast = combined_data$`XB1 Mamedova Forecast`,
XB1_FVD_Forecast = combined_data$`XB1 FVD Forecast`,
                                   XB2 = rep(c(Waleed_Data$XB2, rep(NA, 12))),
XB2_Mamedova_Forecast = combined_data$`XB2 Mamedova Forecast`,
XB2_FVD_Forecast = combined_data$`XB2 FVD Forecast`,
                                   XB3 = rep(c(Waleed_Data$XB3, rep(NA, 12))),
XB3_Mamedova_Forecast = combined_data$`XB3 Mamedova Forecast`,
XB3_FVD_Forecast = combined_data$`XB3 FVD Forecast`,
                                   XB4 = rep(c(Waleed_Data$XB4, rep(NA, 12))),
XB4_Mamedova_Forecast = combined_data$`XB4 Mamedova Forecast`,
XB4_FVD_Forecast = combined_data$`XB4 FVD Forecast`,
                                   XB5 = rep(c(Waleed_Data$XB5, rep(NA, 12))),
XB5_Mamedova_Forecast = combined_data$`XB5 Mamedova Forecast`,
XB5_FVD_Forecast = combined_data$`XB5 FVD Forecast`)

OG_and_Forecasted_df
```

### Creating the Combined Forecasted Data plot

```{r}

# Pivoting the data
long_df2 <- tidyr::pivot_longer(OG_and_Forecasted_df, cols = -Date, names_to = "Variable", 
                                 values_to = "Value")

# Plotting the data
 ggplot(long_df2, aes(x = Date, y = Value, color = Variable)) +
     geom_line() +
     labs(title = "Original + Forecasted Data",
          x = "Years",
          y = "Value") +
     theme_minimal()
 
```
### Goodness of Fit measures
```{r}
#Installing needed package
install.packages("hydroGOF")
```

```{r}
#Calling the needed packages 
library(zoo)
library(hydroGOF)

#Creating the "Goodness of Fit" Data-Set
Goodness_of_Fit_m.df <- data.frame(
XB1_Mamedova_GOF = gof(sim = OG_and_Forecasted_df$XB1_Mamedova_Forecast, obs = OG_and_Forecasted_df$XB1),
XB1_FVD_GOF = gof(sim = OG_and_Forecasted_df$XB1_FVD_Forecast, obs = OG_and_Forecasted_df$XB1),

XB2_Mamedova_GOF = gof(sim = OG_and_Forecasted_df$XB2_Mamedova_Forecast, obs = OG_and_Forecasted_df$XB2),
XB2_FVD_GOF = gof(sim = OG_and_Forecasted_df$XB2_FVD_Forecast, obs = OG_and_Forecasted_df$XB2),

XB3_Mamedova_GOF = gof(sim = OG_and_Forecasted_df$XB3_Mamedova_Forecast, obs = OG_and_Forecasted_df$XB3),
XB3_FVD_GOF = gof(sim = OG_and_Forecasted_df$XB3_FVD_Forecast, obs = OG_and_Forecasted_df$XB3),

XB4_Mamedova_GOF = gof(sim = OG_and_Forecasted_df$XB4_Mamedova_Forecast, obs = OG_and_Forecasted_df$XB4),
XB4_FVD_GOF = gof(sim = OG_and_Forecasted_df$XB4_FVD_Forecast, obs = OG_and_Forecasted_df$XB4),

XB5_Mamedova_GOF = gof(sim = OG_and_Forecasted_df$XB5_Mamedova_Forecast, obs = OG_and_Forecasted_df$XB5),
XB5_FVD_GOF = gof(sim = OG_and_Forecasted_df$XB5_FVD_Forecast, obs = OG_and_Forecasted_df$XB5)
)

# Showing the data set
Goodness_of_Fit_m.df
```


### Write the "Combined Forecasted" & "Goodness of Fit" data-frames to a CSV files
```{r}
write.csv(OG_and_Forecasted_df , file = "Original & 2 ways Forecast.csv", row.names = FALSE)
write.csv(Goodness_of_Fit_m.df , file = "The Goodness of Fit for the forecast.csv", row.names = TRUE)
```
